% !TEX root = ../thesis.tex

\chaptermark{Introduction}
\phantomsection
\addcontentsline{toc}{chapter}{Introduction}

\chapter*{Introduction}

Linear prediction~\cite{Vaidyanathan} is a method used in signal processing to predict future values of a time series
based on past observations. The technique is based on the assumption that the signal can be modeled as a linear
combination of past values and a noise term. Long-term linear prediction refers to the application of this method
to predict values over a longer period of time, such as months or years.\\
\\
It requires a greater amount of data and is more complex than short-term prediction, but can be useful in areas
such as stock market forecasting and weather prediction. The goal of this master thesis is to developing new
algorithms and mathematical models to improve the accuracy of long-term predictions and use signal principles
in sales forecasting. Matlab \footnote{MATLAB is a fourth-generation programming language and numerical analysis
environment. Uses for MATLAB include matrix calculations, developing and running algorithms, creating user interfaces (UI)
and data visualization.} livescript~\cite{livescript} will be used as development environment to validate our
models and machine learning developed approaches.

\section*{Task formulation}

Proposed a mathematical model and algorithms for sales forecasting based on long-term prediction with improved
Levinson - Durbin scheme which should have better performance and accuracy than known linear prediction
mechanism in the sales forecasting.
